### 2016-06-07 paper/icfp16-cameraready/abstract.tex
init paper/icfp16-cameraready
clarify various things relating to language features and inhabitants

\begin{abstract}
  %
  \emph{Static} type errors are a common stumbling block
  for newcomers to typed functional languages.
  %
  We present a \emph{dynamic} approach to explaining type
  errors by generating counterexample witness inputs that
  illustrate \emph{how} an ill-typed program goes wrong.
  %
  First, given an ill-typed function, we symbolically
  execute the body to dynamically synthesize witness values
  that can make the program go wrong.
  We prove that our procedure synthesizes
  \emph{general witnesses} in that if a witness is
  found, then \emph{for all} input types, there exist
  inhabitants that can make the function go wrong.
  %
  Second, we show how to extend the above procedure to
  produce a \emph{reduction graph} that can be used to
  interactively visualize and debug witness executions.
  %
  Third, we evaluate our approach on two data sets
  comprising over 4,500 ill-typed student programs.
  Our technique is able to generate witnesses for
  88\% of the programs, and our reduction graph
  yields small counterexamples for 90\% of the witnesses.
\end{abstract}

>>>>

\begin{abstract}
  %
  \emph{Static} type errors are a common stumbling block
  for newcomers to typed functional languages.
  %
  We present a \emph{dynamic} approach to explaining type
  errors by generating counterexample witness inputs that
  illustrate \emph{how} an ill-typed program goes wrong.
  %
  First, given an ill-typed function, we symbolically
  execute the body to dynamically synthesize witness values
  that can make the program go wrong.
  We prove that our procedure synthesizes
  \emph{general witnesses} in that if a witness is
  found, then \emph{for all} inhabited input types, there exist
  values that can make the function go wrong.
  %
  Second, we show how to extend the above procedure to
  produce a \emph{reduction graph} that can be used to
  interactively visualize and debug witness executions.
  %
  Third, we evaluate our approach on two data sets
  comprising over 4,500 ill-typed student programs.
  Our technique is able to generate witnesses for
  88\% of the programs, and our reduction graph
  yields small counterexamples for 81\% of the witnesses.
\end{abstract}

### 2016-06-07 paper/icfp16-cameraready/evaluation.tex
clarify various things relating to language features and inhabitants

We have implemented a prototype of our search procedure and trace
visualization for the purely functional subset of \ocaml, in a tool called \nanomaly.
In our implementation we instantiated \gensym\ with a simple random
generation of values, which we will show suffices for the
majority of type errors.

>>>>

We have implemented a prototype of our search procedure and trace
visualization for a purely functional subset of \ocaml\
--- with polymorphic types and records, but no modules, objects, or polymorphic variants ---
in a tool called \nanomaly.
%
We treat explicit type signatures, \eg @(x : int)@, as
primitive operations that narrow the type of the wrapped value.
%
In our implementation we instantiated \gensym\ with a simple random
generation of values, which we will show suffices for the majority of
type errors.

### 2016-06-07 paper/icfp16-cameraready/introduction.tex
clarify various things relating to language features and inhabitants

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:semantics}).
%
This problem is surprisingly tricky when we cannot rely on
static type information.
%
In particular, we must avoid the trap of \emph{spurious} inputs
that cause irrelevant problems that would be avoided by picking
values of a different, relevant type.
%
We solve this problem by developing a novel operational semantics
that combines evaluation and type inference.
%
We execute the program with \emph{holes} --- values whose type is
unknown --- as the inputs.
% that corresponds to executing the program with \emph{holes} ---
% values whose type is unknown.
%
A hole remains abstract until the evaluation context tells us what type
it must have, for example the parameters to an addition operation must
both be integers.
%
Our semantics \emph{conservatively} instantiates holes with concrete
values; thereby dynamically inferring the type of the input
until the program goes wrong.
%
We prove that our procedure synthesizes \emph{general witnesses},
which means, intuitively, that if a witness is found for a given
ill-typed function, then, \emph{for all} input types, there exist
inhabitants that can make the function go wrong.

>>>>

\paragraph{1. Finding Witnesses}
Our first contribution is an algorithm for searching for
\emph{witnesses} to type errors, \ie\ inputs that cause a
program to go wrong~(\S~\ref{sec:semantics}).
%
This problem is surprisingly tricky when we cannot rely on
static type information.
%
In particular, we must avoid the trap of \emph{spurious} inputs
that cause irrelevant problems that would be avoided by picking
values of a different, relevant type.
%
We solve this problem by developing a novel operational semantics
that combines evaluation and type inference.
%
We execute the program with \emph{holes} --- values whose type is
unknown --- as the inputs.
% that corresponds to executing the program with \emph{holes} ---
% values whose type is unknown.
%
A hole remains abstract until the evaluation context tells us what type
it must have, for example the parameters to an addition operation must
both be integers.
%
Our semantics \emph{conservatively} instantiates holes with concrete
values; thereby dynamically inferring the type of the input
until the program goes wrong.
%
We prove that our procedure synthesizes \emph{general witnesses}, which
means, intuitively, that if a witness is found for a given ill-typed
function, then, \emph{for all} (inhabited) input types, there exist
values that can make the function go wrong.
